---
title: "R Notebook"
output: html_notebook
---
#loading library

```{r} 
library(nycflights13) 
library(tidyverse)
```


#loading data:
```{r}
flights <- flights
head(flights)
```

`arrange()` works similarly to `filter()` except that instead of selecting rows, it changes their order. It takes a data frame and a set of column names (or more complicated expressions) to order by. If you provide more than one column name, each additional column will be used to break ties in the values of preceding columns:

```{r}
flights %>% 
  arrange(year, desc(month), desc(day), desc(hour), desc(minute)) %>%
  select(dest, carrier, year, month, day, hour, minute)
```
Use `desc()` to re-order by a column in descending order:

```{r}
nycflights13::airlines
```


```{r}
flights %>% 
  arrange(desc(dep_delay))
```


Missing values are always sorted at the end:

```{r}
df <- tibble(x = c(5, 2, NA))


df 
df %>%
  arrange(desc(x))
```
5.3.1 Exercises
1. How could you use `arrange()` to sort all missing values to the start? (Hint: use `is.na()`).

```{r}
df %>%
  filter(!is.na(x))
```

```{r}
head(flights)
```

2. Sort flights to find the most delayed flights. Find the flights that left earliest.

```{r}
flights %>%
  arrange(dep_delay)

```


3. Sort flights to find the fastest (highest speed) flights.

```{r}
flights %>%
  mutate(hours = air_time/60,
         mph = distance/hours) %>%
  arrange(desc(mph))
```


```{r}
nycflights13::planes
```


```{r}
flights %>%
  mutate(hours = air_time/60,
         mph = distance/hours) %>%
  arrange(desc(mph)) %>%
  select(tailnum, mph) %>%
  left_join(nycflights13::planes, by = "tailnum")
```


4. Which flights travelled the farthest? Which travelled the shortest?

```{r}
nycflights13::airports

flights %>%
  arrange(desc(distance)) %>%
  distinct(origin, dest) %>% #unique pairs
  left_join(airports, by=c("dest" = "faa")) 
```



Select columns with `select()`
It’s not uncommon to get datasets with hundreds or even thousands of variables. In this case, the first challenge is often narrowing in on the variables you’re actually interested in. `select()` allows you to rapidly zoom in on a useful subset using operations based on the names of the variables.

`select()` is not terribly useful with the flights data because we only have 19 variables, but you can still get the general idea:

```{r}
# Select columns by name
select(flights, year, month, day)
```


```{r}
# Select all columns between year and day (inclusive)
select(flights, year:day)
```

```{r}
# Select all columns except those from year to day (inclusive)
select(flights, -(year:day))
```

There are a number of helper functions you can use within `select()`:

`starts_with("abc")`: matches names that begin with “abc”.

`ends_with("xyz")`: matches names that end with “xyz”.

`contains("ijk")`: matches names that contain “ijk”.

`matches("(.)\\1")`: selects variables that match a regular expression. This one matches any variables that contain repeated characters. You’ll learn more about regular expressions in strings.

`num_range("x", 1:3)`: matches x1, x2 and x3.

See `?select` for more details.

`select()` can be used to rename variables, but it’s rarely useful because it drops all of the variables not explicitly mentioned. Instead, use `rename()`, which is a variant of `select()` that keeps all the variables that aren’t explicitly mentioned:

Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.
